---
title: "Lab 01 - Hello R"
author: "Anahatt Virk"
date: "6/17/2025"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

The datasaurus_dozen file has 1846 rows and 3 columns. The variables included in the data frame are dataset, x, and y. 

### Exercise 2

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
The correlation between x and y is approximately -0.065, suggesting that there is no linear relationship present between the two variables. However, when visualizing the data we are able to see a clear dinosaur shape, which emphasizes the importance of visualizing the data instead of solely relying on the statistics. 

### Exercise 3

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x=x, y = y)) +
  geom_point()

```

Correlation between `x` and `y` in this dataset:

```{r cor-star}
star_data %>%
  summarize(r = cor(x,y))
```

The correlation between x and y in the star dataset is about -0.063, suggesting that there is no meaningful linear relationship between the x and y coordinates. Similar to Exercise 2, when plotting the data we see a clear star shape plotted, once again demonstrating that relying on summary statistics in isolation can be misleading. 

### Exercise 4

```{r plot-circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x=x, y = y)) +
  geom_point()
```

Correlation between `x` and `y` in this dataset:

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x,y))
```

The correlation between x and y for the circle dataset is -0.068, once again highlighting the importance of data visualization to uncover patterns that statistics alone might miss. 

### Exercise 5

```{r plot-bullseye}
bullseye_data <- datasaurus_dozen %>%
  filter(dataset == "bullseye")

ggplot(data = bullseye_data, mapping = aes(x=x, y = y)) +
  geom_point()
```

Correlation between `x` and `y` in this dataset:

```{r cor-bullseye}
bullseye_data %>%
  summarize(r = cor(x,y))
```

With a correlation coefficient of about -0.069, the bullseye dataset also demonstrates the significance of using both statistics and data visualization hand in hand to uncover patterns in the data. 