---
title: "Lab 01 - Hello R"
author: "Anahatt Virk"
date: "6/17/2025"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
    highlight: tango
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

The datasaurus_dozen file has 1846 rows and 3 columns. The variables included in the data frame are dataset, x, and y. 

### Exercise 2

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
The correlation between x and y is approximately -0.065, suggesting that there is no linear relationship present between the two variables. However, when visualizing the data we are able to see a clear dinosaur shape, which emphasizes the importance of visualizing the data instead of solely relying on the statistics. 

### Exercise 3

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x=x, y = y)) +
  geom_point()

```

Correlation between `x` and `y` in this dataset:

```{r cor-star}
star_data %>%
  summarize(r = cor(x,y))
```

The correlation between x and y in the star dataset is about -0.063, suggesting that there is no meaningful linear relationship between the x and y coordinates. Similar to Exercise 2, when plotting the data we see a clear star shape plotted, once again demonstrating that relying on summary statistics in isolation can be misleading. The obtained value for this exercise is fairly similar to the r for dino. 

### Exercise 4

```{r plot-circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x=x, y = y)) +
  geom_point()
```

Correlation between `x` and `y` in this dataset:

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x,y))
```

The correlation between x and y for the circle dataset is -0.068, again highlighting the importance of data visualization to uncover patterns that statistics alone might miss. This value is also almost identical to the r for dino. 

### Exercise 5

```{r plot-all}

ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```

Correlation between `x` and `y` in all datasets:

```{r cor-all}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(n = 13)
```

As can be seen in the above table, most of the datasets have very similar correlation values. This demonstrates the significance of data visualization. While all these datasets have nearly identical summary statistics, they look very different when plotted, emphasizing the need to rely on both statistics and visualization when completing research. 